Sulfur Dioxide on Water's Surface

### 2011-03-21 Methods.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

Molecular dynamics simulations were all performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 model,\cite{Caldwell1995} and for \suldiox we used the model of Dang, et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010}

All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs. The long axis of each simulation cell (the axis normal to the water surface) was then lengthened to 120\angs, and the systems were further equilibrated for 10 ns. The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry. After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system.

One system involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration. The second system introduced 22 \suldiox~molecules into the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab, on both surfaces, to simulate an added 50 atm of \suldiox~gas pressure. The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The complementary saturated system with 22 \suldiox~molecule in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface. Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.

@@

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations involved an equilibrated water slab as in the surface equilibrated method above. However, for both the neat-water and saturated systems, a single \suldiox~was introduced 20\angs above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water-oxygen 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.

A separate set of SMD simulations were performed similarly, but with the tethering to one of the \suldiox-oxygens. This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering. The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/startingconfigurations.png}
		\caption{Sample starting configurations for the two types of SMD simulations. The neat-\wat~slab simulation introduces a single \suldiox~molecule that is then guided into the surface of the water and further into the bulk (left). The saturated-\wat~slab simulation begins with different configurations of a water system that has been loaded with \suldiox~to saturate the water phase, and also with a high pressure of \suldiox~gas (right). The single \suldiox~(shown at the top) is then steered through the surface region saturated with \suldiox~molecules, and into the water bulk.}
		\label{fig:starting-configurations}
	\end{center}
\end{figure}

>>>>

Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Dang, et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010}

All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs. The long axis of each simulation cell (the axis normal to the water surface) was then lengthened to 120\angs, and the systems were further equilibrated for 10 ns. The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry. After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system with multiple gaseous \suldiox~molecules.

The neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration. The saturated system had 22 \suldiox~molecules in the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab, on both surfaces, to simulate an added 50 atm of \suldiox~gas pressure. The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The saturated system with 22 \suldiox~molecules in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface. Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.

@@

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above. However, in both the neat-water and saturated systems equilibrated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.

A separate set of SMD simulations were performed with tethering to one of the \suldiox-oxygens. This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering. The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/startingconfigurations.png}
		\caption{Sample starting configurations for the two types of SMD simulations. The neat-water slab simulation introduces a single \suldiox~molecule that is then guided into the surface of the water and further into the bulk (left). The saturated slab simulation begins with a water system that has been loaded with \suldiox~to saturate the water phase, and also with a high pressure of \suldiox~gas (right). The single \suldiox~(shown at the top) is then steered through the surface region saturated with \suldiox~molecules, and into the water bulk.}
		\label{fig:starting-configurations}
	\end{center}
\end{figure}

### 2011-03-21 TransitOrientation.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

	The orientation of a \suldiox~molecule throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles}. Only the surface into which the transiting \suldiox~was steered into is shown for the saturated system's analysis. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position 20\angs above the water surface, until the \suldiox~moves to within 10\angs of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. The bisector angle $\theta$, near to the surface becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance$=0$), the bisector is perpendicular in both the neat-water and saturated systems. There are differences between the neat-water and saturated systems, however, in the onset points of the orientational preferences.

  % talk about theta differences
  The $\theta$ distribution clearly shows that the \suldiox~is isotropically oriented in the gas-phase and then orients with its bisector perpendicular to the water surface within a certain distance above the water phase. On approach to the neat-water surface, that transition occurs once just above 10\angs from the surface, and again at 5\angs from the surface and below. The onset of orientation at 10\angs appears to be an artifact of the small data set and the normalization scheme, as the isotropic behavior resumes between 5-10\angs. Below 5\angs, however, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approx. 8\angs above the water surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few\angs within the \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.

  % talk about phi differences
  With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis. This is the case in both systems with one exception in the neat-water system. At the water surface and just below from 0\angs to 5\angs in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$. This corresponds to a \suldiox~orienting more flat to the surface (inclined up to 60 degrees from the surface normal). However, below this 5\angs region, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

>>>>

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles}. Only the surface into which the transiting \suldiox~was steered is used in the analyses. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

@@

	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. The bisector angle $\theta$, near to the surface becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. There are differences between where the orientation transitions from isotropic in the neat-water and saturated systems. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.

  % talk about theta differences
  The $\theta$ distribution clearly shows that the \suldiox~is isotropically oriented in the gas-phase and then orients with its bisector perpendicular to the water surface within a certain distance above the water phase. On approach to the neat-water surface, that transition occurs once just above 10\angs~from the surface, and again at 5\angs~from the surface and below. The onset of a preferred orientation at 10\angs~occurrs at the cutoff length of the simulated intermolecular bonding interactions. Between 5-10\angs~the molecule appears to orient more isotropically. Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approx. 8\angs~above the surface location. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few\angs~within the \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.

  % talk about phi differences
  With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis. This is the case in both systems with one exception in the neat-water system. At the water surface and just below from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$. This corresponds to a \suldiox~orienting more flat to the surface (inclined up to 60 degrees from the surface normal). However, below this 5\angs~region, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
